---
output: github_document
title: "kcHARS"
---

<!-- following ensures the air R code formatter doesn't format this document (since it doesn't support documents other than .R files)  -->

```{r include=FALSE}
# fmt: skip file
```

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start --> 
<!-- badges: end -->

The kcHARS R package provides functions for reading and working with eHARS data.

## Installation 

You can install the development version of kcHARS from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("rlechtenberg/kcHARS")
```

## Reading eHARS data into R 

kcHARS assumes eHARS SAS datasets reside in "R:/Epi/EXTRA SPACE/eHARS/Full
eHARS" (what `ehars_root_dir()` returns). Specifically, it assumes that the
./Archive subdirectory contains folders named 'YYYY-MM-DD', each
containing an extract from a different date, and that each of those contains a
./person_sas7bdat subdirectory containing eHARS' person-based dataset (AKA the
person view) and a ./document_sas7bdat subdirectory containing the
document-based datasets. When you load and attach the kcHARS package using the
`library()` function, R identifies the date of the most recent archived eHARS
extract, stores it in a global option named kcHARS.ehars_dt, and notifies you of
it via the package start-up message. Calls to `ehars_person()` and `ehars_doc()`
will then read the requested data from the associated folder.

```{r example}
library(kcHARS)

# print date of eHARS data being used
getOption("kcHARS.ehars_dt")

# get max document enter_dt
ehars_doc("document", col_select = c("enter_dt")) |>
  dplyr::filter(enter_dt == max(enter_dt)) |>
  dplyr::distinct() |>
  dplyr::mutate(enter_dt_num = ehars_dt.as.Date(enter_dt)) 
```

If you want to use the data from a different/older extract, you can simply
update that kcHARS.ehars_dt global option using `options(kcHARS.ehars_dt =
'<YYYY-MM-DD>')`.

```{r}
options(kcHARS.ehars_dt = "2025-12-30")

# print date of eHARS data being used
getOption("kcHARS.ehars_dt")

# get max document enter_dt
ehars_doc("document", col_select = c("enter_dt")) |>
  dplyr::filter(enter_dt == max(enter_dt)) |>
  dplyr::distinct() |>
  dplyr::mutate(enter_dt_num = ehars_dt.as.Date(enter_dt))
```


<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` -->
<!-- up-to-date. `devtools::build_readme()` is handy for this.  -->
